
\chapter{Introduction\label{ch:intro}}

Since the better part of the modern era, the increasing topological changes in urban environment have caused the human civilization to be subjected to an increasing risk of disasters due to natural or artificial, and/or a combination of the two causes. Hence it has become an ever increasing need to design infrastructure to handle such disasters in the event it may occur. But even more so, the safe evacuation of people and personnel with the premises of the affected infrastructure takes precedence when dealing the necessary mitigation and disaster risk management. The evacuation time of people from a scene of disaster is extremely crucial in the case of an emergency in disaster situation. In order to reduce the time taken for evacuation, better and more robust exit strategy evacuation algorithms are developed which are used to model participating agents for their exit patterns and exit strategies and clock them based on performance, efficiency and practicality. In order to evaluate such parameters, evacuation simulation models have been developed and constantly tested over the years to investigate the emergency egress capabilities of the built environment for a variety of reasons including: difficulty in conducting real evacuation tests (drills); aiding building design and confirming conformity to building regulations; and even determining optimal evacuation routes for the building occupiers \cite{ref1}. Computational models were first introduced in the early 1980s since the inception of computers. During the 90s, academic researchers aimed to improve the capabilities of these models in order to optimize and improve the pathfinding performance of the evacuees and their corresponding movements \cite{ref2}. Over the course of many years, these algorithms have become the industry standard to perform exit evacuation time and performance analysis. The purpose of this thesis is to study the state of the art algorithm and to draw comparisons between micro model agent simulation with respect to its counterpart, the macro model agent simulation.

\input{ch-intro/intro_contributions}
